The non-random assembly of indirect interactions in  \nplant-pollinator networks

### 2022-04-26 Manuscript/Main_text/Main_text_Journal_of_Animal_Ecol.Rmd
edits on text

Ecological processes leave distinct structural imprints on the species interactions that shape the topology of animal-plant mutualistic networks. Detecting how direct and indirect interactions between animals and plants are organized is not trivial since they go beyond pairwise interactions, but may get blurred when considering full network descriptors. Recent work has shown that the network meso-scale can capture this important information. The meso-scale describes network subgraphs representing patterns of direct and indirect interactions between a small number of species (i.e., motifs) that constitute the building blocks of the whole network. Despite the possible implications of network motifs to better capture species interactions, they remain overlooked in natural plant-pollinator networks. By exploring 60 empirical plant-pollinator networks from 18 different studies with wide geographical coverage we show that some motifs are consistently under- or over-represented worldwide, suggesting that the building blocks of plant-pollinator networks are not random. We found a higher proportion of densely connected motifs, which could reflect that neutral effects are the predominant ecological process governing the structure of network motifs. Furthermore, we find that distinct motif positions describing species ecological roles (e.g., generalisation and number of indirect interactions) are occupied by different floral visitor and plant groups on both trophic levels. For example, bees tend to appear in motif positions with a lower number of indirect interactions than the rest of floral visitor groups. Interestingly, the realized combinations of different species groups within a motif can not be retrieved from their joint probability distributions, indicating that group combinations are not random either. Our result highlight the non-random structure of the key building blocks shaping plant-pollinators networks and the association of different floral visitor and plant groups with certain motifs.

>>>>

Ecological processes leave distinct structural imprints on the species interactions that shape the topology of animal-plant mutualistic networks. Detecting how direct and indirect interactions between animals and plants are organized is not trivial since they go beyond pairwise interactions, but may get blurred when considering full network descriptors. Recent work has shown that the network meso-scale can capture this important information. The meso-scale describes network subgraphs representing patterns of direct and indirect interactions between a small number of species (i.e., motifs) that constitute the building blocks of the whole network. Despite the possible implications of network motifs to better capture species interactions, they remain overlooked in natural plant-pollinator networks. By exploring 60 empirical plant-pollinator networks from 18 different studies with wide geographical coverage we show that some motifs are consistently under- or over-represented worldwide, suggesting that the building blocks of plant-pollinator networks are not random. We found a higher proportion of densely connected motifs, which could reflect that neutral effects are the predominant ecological process governing the structure of network motifs. Furthermore, we find that distinct motif positions describing species ecological roles (e.g., generalisation and number of indirect interactions) are occupied by different floral visitor and plant groups on both trophic levels. For example, bees tend to appear in motif positions with a lower number of indirect interactions than the rest of floral visitor groups. Interestingly, the realized combinations of different species groups within a motif can not be retrieved from their joint probability distributions, indicating that group combinations are not random either. Our results highlight the non-random structure of the key building blocks shaping plant-pollinators networks and the association of certain motifs with different groups of floral visitors and plants.
